---
title: "Getting Started with AgentsMCP"
author: "Avisek Banerjee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AgentsMCP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction to AgentsMCP

The AgentsMCP package provides a comprehensive implementation of the Agentic Neural Network Economic Model (ANNEM), featuring heterogeneous AI agents with neural decision-making capabilities, Model Context Protocol (MCP) communication, and dynamic network formation.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("avishekb9/AgentsMCP")
```

```{r setup}
library(AgentsMCP)
library(ggplot2)
library(dplyr)
```

## Quick Start Example

Let's start with a simple ANNEM analysis using a small number of agents:

```{r basic_example, eval=FALSE}
# Set seed for reproducibility
set_annem_seed(42)

# Run a basic ANNEM analysis
results <- run_annem_analysis(
  symbols = c("AAPL", "MSFT"),
  n_agents = 100,
  n_steps = 50,
  save_results = FALSE,
  verbose = FALSE
)

# View summary
annem_summary(results)
```

## Understanding Agent Types

ANNEM implements six different types of heterogeneous agents:

1. **Neural Momentum**: Trend-following agents with neural network enhancement
2. **Contrarian AI**: Mean-reversion agents with AI-based signals  
3. **Fundamentalist ML**: Technical analysis agents with machine learning
4. **Adaptive Noise**: Random strategy agents with adaptive learning
5. **Social Network**: Peer influence and herding behavior agents
6. **Meta Learning**: MAML-inspired strategy adaptation agents

### Creating Individual Agents

```{r agent_creation}
# Create different types of agents
momentum_agent <- create_annem_agent("agent_001", "neural_momentum", 1000000)
contrarian_agent <- create_annem_agent("agent_002", "contrarian_ai", 1000000)

# Display agent properties
cat("Momentum Agent:\n")
cat("- ID:", momentum_agent$id, "\n")
cat("- Type:", momentum_agent$type, "\n")
cat("- Wealth:", scales::dollar(momentum_agent$wealth), "\n")
cat("- Risk Tolerance:", round(momentum_agent$risk_tolerance, 3), "\n\n")

cat("Contrarian Agent:\n")
cat("- ID:", contrarian_agent$id, "\n")
cat("- Type:", contrarian_agent$type, "\n")
cat("- Wealth:", scales::dollar(contrarian_agent$wealth), "\n")
cat("- Risk Tolerance:", round(contrarian_agent$risk_tolerance, 3), "\n")
```

## Working with Market Data

The package can automatically download market data or work with synthetic data:

```{r synthetic_data}
# Generate synthetic market data for testing
synthetic_data <- generate_synthetic_data(
  n_days = 100,
  n_assets = 2,
  annual_return = 0.08,
  annual_volatility = 0.20
)

cat("Generated", nrow(synthetic_data$prices), "days of data for", 
    ncol(synthetic_data$prices), "assets\n")
cat("Assets:", paste(synthetic_data$symbols, collapse = ", "), "\n")

# Plot synthetic price evolution
if (requireNamespace("ggplot2", quietly = TRUE)) {
  price_df <- data.frame(
    day = 1:nrow(synthetic_data$prices),
    price = synthetic_data$prices[, 1]
  )
  
  ggplot(price_df, aes(x = day, y = price)) +
    geom_line(color = "steelblue", size = 1) +
    labs(
      title = "Synthetic Asset Price Evolution",
      x = "Trading Day",
      y = "Price ($)"
    ) +
    theme_minimal()
}
```

## Creating and Running Market Simulations

### Basic Market Creation

```{r market_creation, eval=FALSE}
# Create a market environment
market <- create_annem_market(
  n_agents = 200,
  symbols = c("AAPL", "MSFT", "GOOGL")
)

# Examine the market
cat("Market created with", length(market$agents), "agents\n")
cat("Network density:", round(igraph::edge_density(market$network), 4), "\n")

# Check agent type distribution
agent_types <- sapply(market$agents, function(a) a$type)
type_distribution <- table(agent_types)
print(type_distribution)
```

### Running Simulations

```{r simulation_example, eval=FALSE}
# Run simulation with the market
simulation_results <- market$run_simulation(n_steps = 100, verbose = TRUE)

# Analyze results
cat("Simulation completed with", simulation_results$n_steps, "steps\n")
cat("Agent decisions matrix dimensions:", 
    paste(dim(simulation_results$agent_decisions), collapse = " x "), "\n")
cat("Wealth evolution matrix dimensions:", 
    paste(dim(simulation_results$agent_wealth), collapse = " x "), "\n")
```

## Performance Analysis

### Agent Performance Metrics

```{r performance_analysis, eval=FALSE}
# Analyze agent performance
agent_performance <- market$analyze_agent_performance()

# Summary by agent type
performance_summary <- agent_performance %>%
  group_by(agent_type) %>%
  summarise(
    count = n(),
    avg_return = mean(total_return) * 100,
    median_return = median(total_return) * 100,
    avg_sharpe = mean(sharpe_ratio),
    .groups = 'drop'
  ) %>%
  arrange(desc(avg_return))

print(performance_summary)
```

### Network Analysis

```{r network_analysis, eval=FALSE}
# Analyze network evolution
network_metrics <- market$analyze_network_evolution()

# Display network evolution summary
cat("Network Evolution Summary:\n")
cat("- Initial density:", round(network_metrics$density[1], 4), "\n")
cat("- Final density:", round(tail(network_metrics$density, 1), 4), "\n")
cat("- Average clustering:", round(mean(network_metrics$clustering, na.rm = TRUE), 4), "\n")
```

## Visualization

The package provides comprehensive visualization capabilities:

```{r visualization_example, eval=FALSE}
# Create performance plots
perf_plots <- plot_agent_performance(agent_performance)

# Display distribution plot
print(perf_plots$performance_dist)

# Create network evolution plots
network_plots <- plot_network_evolution(network_metrics)
print(network_plots$evolution)

# Create wealth dynamics plots
wealth_plots <- plot_wealth_dynamics(simulation_results)
print(wealth_plots$by_type)
```

## Model Comparison and Validation

### Benchmark Comparison

```{r benchmarking, eval=FALSE}
# Compare with traditional models
comparison_results <- market$compare_with_benchmarks()

print(comparison_results)

# Visualize comparison
comparison_plots <- plot_model_comparison(comparison_results)
print(comparison_plots$comparison)
```

### Mathematical Framework Validation

```{r validation, eval=FALSE}
# Validate implementation against mathematical framework
validation_results <- validate_annem_framework(market, simulation_results, agent_performance)

# Display validation summary
passed_tests <- sum(validation_results$Validation_Status == "PASS")
total_tests <- nrow(validation_results)

cat("Framework Validation Results:\n")
cat("Tests passed:", passed_tests, "out of", total_tests, "\n")

# Show any failed tests
failed_tests <- validation_results[validation_results$Validation_Status != "PASS", ]
if (nrow(failed_tests) > 0) {
  cat("\nTests requiring attention:\n")
  for (i in 1:nrow(failed_tests)) {
    cat("-", failed_tests$Metric[i], ":", failed_tests$Validation_Status[i], "\n")
  }
}
```

## Advanced Configuration

### Custom Agent Distributions

```{r custom_config}
# Get default configuration
config <- get_annem_config()

cat("Default Agent Types:\n")
for (i in 1:length(config$agent_types)) {
  cat("-", config$agent_types[i], ":", 
      scales::percent(config$agent_distribution[i]), "\n")
}

cat("\nDefault Parameters:\n")
cat("- Default agents:", config$default_n_agents, "\n")
cat("- Default steps:", config$default_n_steps, "\n")
cat("- Risk-free rate:", scales::percent(config$risk_free_rate), "\n")
```

### Performance Optimization

For large-scale simulations, consider these optimization strategies:

```{r optimization_tips, eval=FALSE}
# For testing and development
quick_results <- run_annem_analysis(
  symbols = c("AAPL"),
  n_agents = 50,
  n_steps = 25,
  save_results = FALSE
)

# For production analysis
production_results <- run_annem_analysis(
  symbols = c("AAPL", "MSFT", "GOOGL", "TSLA", "NVDA"),
  n_agents = 1000,
  n_steps = 250,
  save_results = TRUE,
  output_dir = "annem_production_results"
)
```

## Error Handling and Troubleshooting

### Common Issues and Solutions

1. **Internet Connection**: If market data download fails, use synthetic data:

```{r error_handling, eval=FALSE}
# Fallback to synthetic data
tryCatch({
  real_data <- load_market_data(c("AAPL", "MSFT"))
}, error = function(e) {
  cat("Market data download failed, using synthetic data\n")
  real_data <- generate_synthetic_data(n_days = 252, n_assets = 2)
})
```

2. **Memory Issues**: Reduce simulation size for large analyses:

```{r memory_handling, eval=FALSE}
# Check available memory and adjust parameters accordingly
memory_limit <- as.numeric(gsub("[^0-9.]", "", memory.size(max = TRUE)))

if (memory_limit < 8000) {  # Less than 8GB
  n_agents <- 500
  n_steps <- 100
} else {
  n_agents <- 1000
  n_steps <- 250
}

cat("Using", n_agents, "agents and", n_steps, "steps based on available memory\n")
```

## Next Steps

- Explore the **Advanced Modeling** vignette for sophisticated analyses
- Check out the **Network Analysis** vignette for detailed network studies
- Review the **Visualization Guide** for comprehensive plotting options
- Visit the package documentation for complete function references

## References

1. Farmer, J. D., & Foley, D. (2009). The economy needs agent-based modelling. *Nature*, 460(7256), 685-686.
2. Jackson, M. O. (2008). *Social and economic networks*. Princeton University Press.
3. Billio, M., Getmansky, M., Lo, A. W., & Pelizzon, L. (2012). Econometric measures of connectedness and systemic risk in the finance and insurance sectors. *Journal of Financial Economics*, 104(3), 535-559.